---
title: "Driving the sidebar primitive"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Driving the sidebar primitive}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

`blockr.ui` ships a Bootstrap-free slide-in panel that mirrors
`shiny::showModal()` / `shiny::removeModal()` in spirit. Mount the
panel once per page with `sidebar_ui()`, then drive its content from
any `moduleServer()` body using `show_sidebar()` and `hide_sidebar()`.

## Minimal example

```{r}
library(shiny)
library(blockr.ui)

ui <- fluidPage(
  actionButton("open", "Open sidebar"),
  sidebar_ui("main_sidebar")
)

server <- function(input, output, session) {
  observeEvent(input$open, {
    show_sidebar(
      "main_sidebar",
      title = "Add a new block",
      ui = tagList(
        textInput("block_name", "Name"),
        selectInput("block_kind", "Type", c("data", "transform", "plot")),
        actionButton("confirm", "Add", class = "btn-primary")
      )
    )
  })

  observeEvent(input$confirm, {
    # validate, mutate state, ...
    hide_sidebar("main_sidebar")
  })
}

shinyApp(ui, server)
```

The inputs inside the body (`input$block_name`, `input$block_kind`,
`input$confirm`) live in the calling session — exactly as they would
inside a `modalDialog()`. No nested `moduleServer()` is required.

## Observing user-driven dismissals

Each `sidebar_ui()` call also registers a Shiny input under the same
id, exposing two logical fields:

- `input[[id]]$open` — `TRUE` while the panel is open, `FALSE` once
  the user closes it (Esc, X button) or the server calls
  `hide_sidebar()`.
- `input[[id]]$pinned` — `TRUE` while the user has the panel pinned.

```{r}
observe({
  state <- input$main_sidebar
  req(state)
  cat("open:", state$open, " pinned:", state$pinned, "\n")
})
```

## One id per concern

`show_sidebar()` replaces the panel's body and title in place, and pin
state is bound to the panel element. Re-showing the same panel with
fresh content from the same caller (the "chain after confirm" flow,
which `keep_or_hide_sidebar()` builds on) is the supported pattern.

Two unrelated concerns - say a "Settings" panel and an "Add block"
form - should mount two `sidebar_ui()` panels with distinct ids. Sharing
one id between unrelated callers means the second `show_sidebar()` will
silently overwrite the first, including any pinned content the user
expected to stay put.

## Auto-open on an empty board

A common renderer pattern is to open the sidebar with a "first block"
hint when the session starts on an empty board, while still respecting
a later user dismissal. The recipe is built on top of `show_sidebar()`
and `input[[id]]$open` — no special opt-in is needed on `sidebar_ui()`.

```{r}
empty_board_renderer <- function(id, board) {
  moduleServer(id, function(input, output, session) {
    auto_opened <- reactiveVal(FALSE)

    observe({
      if (auto_opened()) return()
      if (length(blockr.core::board_blocks(board)) > 0L) return()

      show_sidebar(
        session$ns("main_sidebar"),
        title = "Add your first block",
        ui = tagList(
          p("Pick a block to get started."),
          actionButton(session$ns("add_dataset"), "Dataset block"),
          actionButton(session$ns("add_import"), "Import data")
        )
      )
      auto_opened(TRUE)
    })

    # Once the user dismisses the hint, input[[id]]$open flips to FALSE
    # and the auto_opened flag prevents this observer from re-opening.
  })
}
```

Push-mode panels reflow the page content via padding on `<html>`,
driven by `--blockr-sidebar-width-left` and
`--blockr-sidebar-width-right` custom properties (both nonzero when
both sides are open). The legacy `--blockr-sidebar-width` is also
set to the open side's width when a single panel is open, so
existing consumers of the form
`margin-right: var(--blockr-sidebar-width, 0px)` keep working.
